### 2013-10-03 Paper_Markdown.Rmd
Some major changes to the coding, mostly adding echo, message and warning blocks to the Markdown R code sections.  Some revisions on the figures.  New table & text to accompany the CCA variance analysis.
Most importantly, I re-ran the MAT runs with fewer values.  It looks like the small intervals we use aren't really all that useful from 0 - 0.2, so I ditched them (hence the change to `bootstrap_mat.R`).  They're not useful for WAPLS or WA but I haven't implemented that yet.  That's why the Figure right before the Discussion has a weird dark/light pattern for WA and WAPLS.  New files come fro

```{r}
library(gridExtra)

@@

```{r SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}
map <- map_data('world')

@@

```{r MinMaxDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations'}

@@

```
**Figure X**.  *Distribution of squared-chord dissimilarity values.  The 95% confidence interval for the minumum spanning distance in the North American Modern Pollen Database is 0.10 - 0.48.  We choose 0.48 as our cut-off for 'non-analogue' distances in the database.*

```{r SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores'}

@@

```
**Figure 2**. *Location of non-analogue points in the modern pollen data, presented within a non-metric multidimensional scaling ordination (Panel 1) and geographically.  Interestingly, although a number of geographically isolated points exist in the dataset, the majority of non-analogue points exist within the most dense region of sampling, indicating that non-analogue sites are not ecologically stratified.*

@@

```{r echo=FALSE, message=FALSE, warning=FALSE}

#  Sakari, can you take a look at this and let me know what you think?
pol <- mod_pol[,7:ncol(mod_pol)]
clim <- climate[,4:ncol(climate)]

@@

names(clim.ratio) <- colnames(clim)
best.clim <- clim.ratio[order(clim.ratio, decreasing = TRUE)]

@@

Key variables based on the $latex \frac {\lambda_1} {\lambda_2} $ using pCCA appear to be climate variables relating to mean monthly percent summer sunshine.  These variables are likely to be variables that are combinations of temperature and preciptiation variables, however this value is not linear, nor is it continuous, meaning that .  While Telford & Birks (2009) have criticised sunshine reconstructions (Freshette et al., 2008) in the past (and perhaps with good reason), the strength of the sunshine signal in the climate data is worth noting, but the limitations of CCA may impact our ability to detect and accurately assess its importance.

@@

Once the climate variables of interest were established we began by bootstrapping the models for these climate variables.

@@

na.samples <- data.frame(site = rep(climate[,1], ncol(wa.res$bias)),
                         size = as.vector(wa.res$sample_size),
                         distance = rep(vals, each = nrow(wa.res$sample_size)),
                         alpha.val = rep(diff(c(-.5,vals)), each = nrow(wa.res$sample_size))+0.005)

@@

ggplot(na.samples, aes(x=distance, y=size)) + 
  geom_path(aes(group=site), alpha=0.001) +
  xlab('Analogue Distance') + ylab('Available Samples') +
  theme(text = element_text(size=24, family='serif')) +
  theme_bw()

@@

load('data/mat.res.RData')
load('data/wa.res.RData')
load('data/wapls.res.RData')

model.out <- data.frame(site   = rep(rep(climate[,1], ncol(mat.res$bias)),3),
                        preds  = c(as.vector(mat.res$mean_prediction),
                                   as.vector(wa.res$mean_prediction),
                                   as.vector(wapls.res$mean_prediction)),
                        dissim = rep(rep(vals, each=nrow(mat.res$bias)),3),
                        model  = rep(c('MAT', 'WA', 'WAPLS'), each = length(mat.res$bias)))

model.out <- model.out[!colSums(is.na(model.out)) > 0,]

ggplot(model.out, aes(x = dissim, y = preds, group = site)) +
  geom_path(alpha = 0.01) +
  theme_bw() +
  facet_wrap(~model, scales='free_y') +
  scale_x_continuous(expand=c(0,0))

>>>>

```{r LoadData, echo=FALSE, warning=FALSE, message=FALSE}
library(gridExtra)

@@

```{r SiteLocations, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}
map <- map_data('world')

@@

```{r MinMaxDissim, fig.width=7, fig.height=6, fig.cap='Figure: North American site locations', echo=FALSE, message=FALSE, warning=FALSE}

@@

```
**Figure 2**.  *Distribution of squared-chord dissimilarity values.  The 95% confidence interval for the minumum spanning distance in the North American Modern Pollen Database is 0.10 - 0.48.  We choose 0.48 as our cut-off for 'non-analogue' distances in the database.*

```{r SiteMDS, fig.width=7, fig.height=6, fig.cap='Figure: North American site MDS scores', echo=FALSE, message=FALSE, warning=FALSE}

@@

```
**Figure 3**. *Location of non-analogue points in the modern pollen data, presented within a non-metric multidimensional scaling ordination (Panel 1) and geographically.  Interestingly, although a number of geographically isolated points exist in the dataset, the majority of non-analogue points exist within the most dense region of sampling, indicating that non-analogue sites are not ecologically stratified.*

@@

```{r message=FALSE, warning=FALSE}

#  Sakari, can you take a look at this and let me know what you think?  Am I doing this right, and what do you make of these results?  Are we simply using too large a dataset?

pol <- mod_pol[,7:ncol(mod_pol)]
clim <- climate[,4:ncol(climate)]

@@

names(clim.ratio) <- colnames(clim)
best.clim <- round(clim.ratio[order(clim.ratio, decreasing = TRUE)]*100, 0)

@@

Variable | Percent Variability
-------- | -------------------
`r names(best.clim)[1]` | `r best.clim[1]`
`r names(best.clim)[2]` | `r best.clim[2]`
`r names(best.clim)[3]` | `r best.clim[3]`
`r names(best.clim)[4]` | `r best.clim[4]`
`r names(best.clim)[5]` | `r best.clim[5]`
`r names(best.clim)[6]` | `r best.clim[6]`
`r names(best.clim)[7]` | `r best.clim[7]`

Key variables based on the $latex \frac {\lambda_1} {\lambda_2} $ using pCCA appear to be climate variables relating to mean monthly percent summer sunshine (variables preceded by an 's').  These variables are likely to be variables that are combinations of temperature and preciptiation variables, however this value is not linear, nor is it continuous, meaning that .  While Telford & Birks (2009) have criticised sunshine reconstructions (Freshette et al., 2008) in the past (and perhaps with good reason), the strength of the sunshine signal in the climate data is worth noting, but the limitations of CCA may impact our ability to detect and accurately assess its importance.

@@

Once the climate variables of interest were established we began by bootstrapping the models for these climate variables.  From the total pool of samples we generate a calibration dataset by sampling with replacement.

@@

na.samples <- data.frame(site = rep(climate[,1], ncol(mat.res$bias)),
                         size = as.vector(mat.res$sample_size),
                         distance = rep(seq(0, 1, by=0.01), each = nrow(mat.res$sample_size)))

@@

loss <- ggplot(na.samples, aes(x=distance, y=size)) + 
  geom_path(aes(group=site), alpha=0.01) +
  xlab('Analogue Distance') + ylab('Available Samples') +
  theme_bw() +
  theme(text = element_text(size=24, family='serif'),
        axis.text = element_text(size=14, family='serif')) +
  scale_x_continuous(expand = c(0,0)) +
  scale_y_continuous(expand=c(0,0)) +
  geom_segment(aes(x=0.48, xend = 0.48, y = min(mat.res$sample_size, na.rm=TRUE), yend = nrow(climate)), size=1, color='red', linetype=2)

loc.loss <- data.frame(lat =  climate$LATDD,
                       long = climate$LONDD,
                       loss = 1 - mat.res$sample_size[,49] / nrow(climate))

loss.map <-  ggplot(data = data.frame(map), aes(long, lat)) + 
  geom_path(aes(group=group), color='black') +
  geom_point(data = loc.loss, 
             aes(x = long, y = lat, color = loss), alpha=0.5) +
   theme_bw() +
  theme(text = element_text(size=24, family='serif'),
        axis.text = element_text(size=14, family='serif')) +
  coord_map(projection='albers', lat0=30, lat1=50, xlim=c(-140, -60), ylim=c(26, 75)) +
  scale_color_continuous(low = '#F08080', high='#4169E1') +
  xlab('') + ylab('')

grid.arrange(loss, loss.map, ncol=2)

@@

mat.df <- data.frame(site = rep(climate[,1], ncol(mat.res$bias)),
                     preds = as.vector(mat.res$mean_prediction),
                     dissim = rep(seq(0, 1, by=0.01),each = nrow(clim)),
                     model = 'MAT', 
                     init = rep(climate[,10], ncol(mat.res$bias)))

model.out <- data.frame(site   = rep(rep(climate[,1], ncol(wa.res$bias)),2),
                        preds  = c(as.vector(wa.res$mean_prediction),
                                   as.vector(wapls.res$mean_prediction)),
                        dissim = rep(rep(vals, each=nrow(wa.res$bias)),2),
                        model  = rep(c('WA', 'WAPLS'), each = length(wa.res$bias)),
                        init = rep(climate[,10], ncol(wa.res$bias)))

#  Temporary fix until we re-run wa.re and wapls.res
model.out <- rbind(mat.df, model.out)
model.out <- model.out[!colSums(is.na(model.out)) > 0,]

#  This plot takes a really long time to plot.

ggplot(model.out, aes(x = dissim, y = preds, group = site, color = init)) +
  geom_path(alpha = 0.05) +
  theme_bw() +
  facet_wrap(~model, scales='free_y') +
  scale_x_continuous(expand=c(0,0)) +
  scale_color_continuous(low='#F08080', high='#4169E1')
